Aquifer depletion exacerbates agricultural drought losses in the US High Plains

### 2023-08-30 Codes/A2_state_fe.rmd
updates on the manuscript

```{r}
#++++++++++++++++++++++++++++++++++++
#+ Formula
#++++++++++++++++++++++++++++++++++++
formula_text <- "acres_ratio ~
    s(sat, k = 5, m = 2) + s(balance_avg, k = 5, m = 2) +
    sandtotal_r + silttotal_r + awc_r +
    factor(year)"

@@

results <-
  all_results[, .(crop, data)] %>%
  expand_grid_df(., formula_data) %>%
  rowwise() %>%
  dplyr::mutate(ir_share_data = list(
    data[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
      .[, balance_avg := mean(balance), by = sc_code]
  )) %>%
  dplyr::mutate(ir_share_res = list(
    gam(
      formula,
      data = ir_share_data_corn,
      family = binomial
    )
  )) %>%
  dplyr::mutate(share_pred_data = list(
    CJ(
      sat = seq(
        min(ir_share_data$sat),
        max(ir_share_data$sat),
        length = 100
      ),
      balance_avg = mean(ir_share_data$balance_avg),
      awc_r = mean(ir_share_data$awc_r),
      sandtotal_r = mean(ir_share_data$sandtotal_r),
      silttotal_r = mean(ir_share_data$silttotal_r),
      state_year = "Nebraska_2009",
      state_name = "Nebraska",
      year = 2009
    )
  )) %>%
  dplyr::mutate(ir_share_hat_data = list(
    predict(
      ir_share_res,
      newdata = share_pred_data,
      type = "response",
      se.fit = TRUE
    )
  )) %>%
  dplyr::mutate(share_pred_data = list(
    dplyr::mutate(
      share_pred_data,
      ir_share_hat = ir_share_hat_data$fit,
      ir_share_hat_se = ir_share_hat_data$se.fit
    )
  )) %>%
  dplyr::select(crop, type, share_pred_data) %>%
  unnest(cols = c(share_pred_data)) %>%
  data.table()
```

```{r}
g_share_comp <-
  results[sat <= 500, ] %>%
  .[, type := factor(type, levels = c("Year FE", "State and Year FE", "State-Year FE"))] %>%
  ggplot() +
  geom_line(aes(y = ir_share_hat, x = sat)) +
  geom_ribbon(
    aes(
      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
      x = sat
    ),
    alpha = 0.4
  ) +
  ylim(0.25, 1) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_x_continuous(breaks = seq(10, 160, by = 20)) +
  xlab("Saturated Thickness (meter)") +
  ylab("Share of Irrigated Acres") +
  facet_grid(type ~ crop) +
  theme_fig +
  theme(
    strip.text = element_text(face = "bold"),
    strip.background = element_blank()
  )

>>>>

```{r}
#++++++++++++++++++++++++++++++++++++
#+ Formula
#++++++++++++++++++++++++++++++++++++
formula_text <- "acres_ratio ~
    s(sat, k = 4, m = 2) + s(balance_avg, k = 4, m = 2) +
    sandtotal_r + silttotal_r + awc_r +
    factor(year)"

@@

# ggplot(results$ir_share_data[[1]]) +
#   geom_point(aes(y = acres_ratio, x = sat))

results <-
  all_results[, .(crop, data)] %>%
  expand_grid_df(., formula_data) %>%
  rowwise() %>%
  dplyr::mutate(ir_share_data = list(
    data %>%
      .[, comp := sum(ir == "nir" | ir == "ir"), by = .(sc_code, year)] %>%
      .[comp == 2, ] %>%
      .[, acres_ratio := acres / sum(acres), by = .(sc_code, year)] %>%
      .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
      .[, balance_avg := mean(balance), by = sc_code] %>%
      .[sat < 180, ]
  )) %>%
  dplyr::mutate(ir_share_res = list(
    gam(
      formula,
      data = ir_share_data,
      family = binomial
    )
  )) %>%
  dplyr::mutate(share_pred_data = list(
    CJ(
      sat = seq(
        min(ir_share_data$sat),
        max(ir_share_data$sat),
        length = 100
      ),
      balance_avg = mean(ir_share_data$balance_avg),
      awc_r = mean(ir_share_data$awc_r),
      sandtotal_r = mean(ir_share_data$sandtotal_r),
      silttotal_r = mean(ir_share_data$silttotal_r),
      state_year = "Nebraska_2009",
      state_name = "Nebraska",
      year = 2009
    )
  )) %>%
  dplyr::mutate(ir_share_hat_data = list(
    predict(
      ir_share_res,
      newdata = share_pred_data,
      type = "response",
      se.fit = TRUE
    )
  )) %>%
  dplyr::mutate(share_pred_data = list(
    dplyr::mutate(
      share_pred_data,
      ir_share_hat = ir_share_hat_data$fit,
      ir_share_hat_se = ir_share_hat_data$se.fit
    )
  )) %>%
  dplyr::select(crop, type, share_pred_data) %>%
  unnest(cols = c(share_pred_data)) %>%
  data.table()
```

```{r}
g_share_comp <-
  results[sat <= 150, ] %>%
  .[, crop := case_when(
    crop == "corn" ~ "Corn",
    crop == "soy" ~ "Soybean"
  )] %>%
  .[, type := factor(type, levels = c("Year FE", "State and Year FE", "State-Year FE"))] %>%
  ggplot() +
  geom_line(aes(y = ir_share_hat, x = sat)) +
  geom_ribbon(
    aes(
      ymin = ir_share_hat - 1.96 * ir_share_hat_se,
      ymax = ir_share_hat + 1.96 * ir_share_hat_se,
      x = sat
    ),
    alpha = 0.4
  ) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), limits = c(0.4, 1)) +
  scale_x_continuous(breaks = seq(10, 150, by = 20)) +
  xlab("Saturated Thickness (meter)") +
  ylab("Share of Irrigated Acres") +
  facet_grid(type ~ crop) +
  theme_fig +
  theme(
    strip.text = element_text(face = "bold"),
    strip.background = element_blank()
  )

### 2023-08-30 Writing/NatureWater/2ndRoundRevision/manuscript.tex
updates on the manuscript

{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_yield_intensive}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_ir_yield_dif}

@@

Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer. Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean. As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.86 (86\%) for an average county-level aquifer thickness of 130m to 0.61 (61\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings). Reductions in the share of production area under irrigation is also evident for soybean although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}.

@@

{\centering \includegraphics[width=5in,height=500px,]{Figures/g_share}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_total_impact}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_avg_yield_dif}

@@

{\centering \includegraphics[width=750px,height=6in,]{Figures/g_y_wd_all}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_map}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_yield_with_conf_corn}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_yield_with_conf_soy}

@@

{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_avg_yield}

@@

{\centering \includegraphics[width=6.5in,height=500px,]{Figures/g_share_dif}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_variation}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_share_comp}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_basis}

@@

{\centering \includegraphics[width=6in,height=500px,]{Figures/g_il_y}

>>>>

{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_yield_intensive}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_ir_yield_dif}

@@

Figure \ref{fig:ir-share} shows the estimated relationship between aquifer thickness and the share of production area that is irrigated for corn and soybean after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer. Figure \ref{fig:ir-share} demonstrates that aquifer thickness has meaningfully large impacts on the share of irrigated production for both corn and soybean. As aquifer thickness declines, the share of irrigated acres for corn decreases from around 0.90 (90\%) for an average county-level aquifer thickness of 100m to 0.72 (72\%) for an average county-level aquifer thickness of 10m (See Appendix \ref{test-dif-share} for formal statistical testings). Reductions in the share of production area under irrigation is also evident for soybean although the effect is less pronounced than for corn, potentially due to soybean's lower water requirements and susceptibility to drought relative to corn \citep{zipper2016drought, ruess2022irrigation}.

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_share}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_total_impact}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_avg_yield_dif}

@@

{\centering \includegraphics[width=750px,height=6in,]{../../Figures/g_y_wd_all}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_map}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_yield_with_conf_corn}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_yield_with_conf_soy}

@@

{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_avg_yield}

@@

{\centering \includegraphics[width=6.5in,height=500px,]{../../Figures/g_share_dif}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_variation}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_share_comp}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_basis}

@@

{\centering \includegraphics[width=6in,height=500px,]{../../Figures/g_il_y}
